Indel Historian: fast and accurate ancestral sequence reconstruction

### 2016-10-18 doc/main.tex
writeup

\begin{table}
  \begin{tabular}{r|r}
    & RMS error \\
    \hline
Indel Historian & \\
PRANK & \\
ProbCons & \\
MUSCLE & \\
  \end{tabular}
  \caption{
    Comparison of Indel Historian to other alignment programs on a simulation benchmark
    using datasets generated by indel-Seq-Gen \citep{StropeEtAl2009}
    with parameters mimicking an alignment of HIV GP120 domains.
  }
\end{table}

@@

We then simulated 100 different alignments under these conditions, using indel-Seq-Gen, and attempted to estimate indel rates
(a) from the true alignment; (b) using alignments or reconstructions created by ProbCons, MUSCLE, PRANK, and Indel Historian.
In each case, we estimated insertion and deletion rates $(\hat{\lambda},\hat{\mu})$ and computed relative errors $(\frac{\hat{\lambda}-\lambda}{\lambda},\frac{\hat{\mu}-\mu}{\mu})$
compared to the true indel rates ($\lambda = \mu = \log(1-p)$).
The RMS (root mean square) of these errors are reported in Table~1.

>>>>

\begin{table}
  \begin{tabular}{r|r}
    & RMS error \\
    \hline
True alignment & \\
Indel Historian & \\
PRANK & \\
ProbCons & \\
MUSCLE & \\
  \end{tabular}
  \caption{
    Comparison of Indel Historian to other alignment programs on a simulation benchmark
    using datasets generated by indel-Seq-Gen \citep{StropeEtAl2009}
    with parameters mimicking an alignment of HIV GP120 domains.
  }
\end{table}

@@

We then simulated 100 different alignments under these conditions, using indel-Seq-Gen, and attempted to estimate indel rates
(a) from the true alignment; (b) using alignments or reconstructions created by ProbCons, MUSCLE, PRANK, and Indel Historian.
In each case, we estimated insertion and deletion rates $(\hat{\lambda},\hat{\mu})$ and computed relative errors $(\frac{\hat{\lambda}-\lambda}{\lambda},\frac{\hat{\mu}-\mu}{\mu})$
compared to the true indel rates ($\lambda = \mu = \log(1-p)$).

The RMS (root mean square) of these errors are reported in Table~1.
Note that even perfect knowledge of the true alignment does not guarantee perfect reconstruction of the underlying indel rate parameters,
due to the possibility of indel events that are not directly observable (for example, when multiple indel events overlap on the same branch,
they will generally be conflated into a single event by the inference algorithm).
